HBCC - DLPFC - FEATURECOUNTS - GRCh38 - CQN - Sparse Model

### 2019-12-19 cmc-rnaseq-de/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd
point to updated metadata file

# Get RNASeq QCmetadata
METADATA_QC_DLPFC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_DLPFC_ID
metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 13) %>%
  dplyr::select(Individual_ID, Institution, Reported_Gender, Sex, Ethnicity, Age_of_Death, `PMI_(in_hours)`, Dx,
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned',
                                      'rnaSeq_report:Transcripts_Detected')) %>% 
  dplyr::filter(Institution %in% c("NIMH-HBCC"))

## Mapping file for misnamed HBCC samples 
maps <- downloadFile("syn17115622") %>% 
  dplyr::filter()

@@

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>% 
  dplyr::rename(PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Reported_Gender,Sex, Ethnicity, Age_of_Death, PMI, Dx, 
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))

# Map to HBCC Sample Ids
METADATA <- dplyr::select(maps, ID, RNA_ID) %>% 
  dplyr::right_join(METADATA, by = c("RNA_ID" = "SampleID")) %>% 
  dplyr::mutate(SampleID = dplyr::coalesce(ID, RNA_ID))

@@

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

>>>>

# Get RNASeq QCmetadata
METADATA_QC_DLPFC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_DLPFC_ID
metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 15) %>%
  dplyr::select(`Individual_ID`, Institution, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx,
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned',
                                      'rnaSeq_report:Transcripts_Detected')) %>% 
  dplyr::filter(Institution %in% c("NIMH-HBCC")) %>% 
  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("DLPFC"))

@@

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>% 
  dplyr::rename(PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))

@@

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

qc <- readRDS(synGet("syn21420622")$path)

```
